---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```

# baggr

-- This is a prototype version of a tool that's in development; we plan to release 1.0 version in October 2018 --

## Overview

Baggr (pronounced as _bagger_ or _badger_) is a meta-analysis package using hierarchical Bayesian models in Stan. Baggr is intended to be user-friendly and transparent so that it's easier to understand the models you're building and criticise them.

## Basic use case

`baggr` is designed to work well with both individual-level ("full") and summarised ("group") data on treatment effect

```{r, eval = FALSE}
library(baggr)
df_pooled <- data.frame("tau" = c(1, -1, .5, -.5, .7, -.7, 1.3, -1.3), 
                        "se" = rep(1, 8))
bgfit <- baggr(df_pooled, pooling = "full")
```

Flexible plotting methods are included, together with comparison of multiple models.

## Realistic use cases

Try `vignette('baggr')` for example of meta-analysis workflow with `baggr`.
