---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
options(tibble.print_min = 5, tibble.print_max = 5)
```



# baggr

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/wwiecek/baggr.svg?branch=cran)](https://travis-ci.org/wwiecek/baggr)
[![codecov](https://codecov.io/gh/wwiecek/baggr/branch/master/graph/badge.svg)](https://codecov.io/gh/wwiecek/baggr)
<!-- badges: end -->

This is _baggr_, an [R package](https://www.r-project.org/) for Bayesian meta-analysis using [Stan](https://mc-stan.org/). _Baggr_ is intended to be user-friendly and transparent so that it's easier to understand the models you are building and criticise them. 

_Baggr_ provides a suite of Bayesian aggregation models that work with both summary data and full data sets, to synthesise evidence collected from different groups, contexts or time periods. The `baggr()` command automatically detects the data type and, by default, fits Rubin's (1981) partial pooling model with weakly informative priors by calling Stan under the hood to carry out Bayesian inference. Modelling of variances or quantiles, standardisation and transformation of data is also possible.

The current version (v0.1, June 2019) is a stable prototype of a tool that's in active development so we are counting on your feedback.




## Installation

Before starting, `baggr` will not work if you don't have RStan. In that case, please follow [the installation instructions for RStan](https://github.com/stan-dev/rstan/wiki/RStan-Getting-Started).

You can install the most up-to-date version of `baggr` directly from GitHub; for this, you will need the `devtools` package.

```{r, eval = FALSE}
# compilation of models should take 5-10 minutes
devtools::install_github("wwiecek/baggr", 
                         build_vignettes = TRUE,
                         build_opts = c("--no-resave-data", "--no-manual"))
```



## Basic use case

`baggr` is designed to work well with both individual-level ("full") and aggregate/summary ("group") data on treatment effect. In basic cases only the summary information on treatment effects (such as means and their standard errors) is needed. Data are always specified in a single data frame and the same `baggr()` function is used for different models. 

For the "standard" cases of modelling means, the appropriate model is detected from the shape of data.

```{r, eval = FALSE}
library(baggr)
df_pooled <- data.frame("tau" = c(28,8,-3,7,-1,1,18,12),
                        "se"  = c(15,10,16,11,9,11,10,18))
bg <- baggr(df_pooled, pooling = "partial")
```

You can specify the model type from several choices, the pooling type (`"none"`, `"partial"` or `"full"`), and certain aspects of the priors, as well as other options for data preparation, prediction and more. You can access the underlying `stanfit` object through `bg$fit`.

Flexible plotting methods are included, together with an automatic comparison of multiple models (e.g. comparing no, partial and full pooling) through `baggr_compare()` command. Various statistics can be calculated: in particular, the `loocv()` command automatically performs leave-one-group-out cross-validation, allowing us to compare and select models. 

Try `vignette('baggr')` for an overview of these functions and an example of meta-analysis workflow with `baggr`. 



## Current and future releases

Included in baggr v0.1:

* Hierarchical models for continuous outcomes
* Meta-analysis specific summaries and plots
* Compatibility with `rstan` and `bayesplot` features
* Both full and aggregate data sets can be used
* Automatic choice of priors
* Automatic calculation of pooling metrics
* Cross-validation

v0.2 will follow soon, enabling more models and more features such as:

* Modelling of quantiles and SE's
* Modelling of log-normal data
* Automatic standardisation of variables
* New way to specify priors
* Leave-one-out cross-validation
